---
title: "Simulation study of the continious versus disceret disease mapping model"
author: "Kamila Kazimierska, Paula Moraga"
date: "December 2020"
output:
  pdf_document: default
  html_document: default
bibliography: references.bib
---
```{r setup, include=FALSE}
lap_data <- "/Users/kazimikk/Dropbox/Paula/Project/"

library(ggplot2)
library(INLA)
library(spdep) # poly2nb 
library(sp)  # over

library("RColorBrewer") # pallete
library(matrixStats) # rowMeans, rowQuantiles, etc
library(data.table) # between CI 
library(knitr) # table 

library(rgeoboundaries) # gb_adm
library(sf)
library(raster) # to read the population object

library(dplyr) # left_join
library(maptools) # unionSpatialPolygon
library(fields) # image 

library(miceadds) # source.all
library(RandomFields) # GF
```

# Simulation study of the continious versus discrete disease mapping models

# Objective 

The point-referenced dataset is measured at known locations. In spatial epidemiology, disease counts are the point-referenced data, but they are aggregated to the health districts or municipalities due to confidentiality reasons. In other words, the continuous process of the disease counts is discretized. 
In the simulation study, we simulate the continuous disease risk surface and counts. Continuous counts means the counts for the fine grid. We compare the performance of the BYM2 model and the model based on the SPDE for the simulated data aggregated into two different administrative levels. 
 

# References 

Additional references to the one specified in the data analysis are:

* [@wang2019disease], simulation of the disease counts using coherent generative model. 
* [@abrahamsen1997review], Gaussian random fields and correlation functions. 
* [@guttorp2006studies], history of the Matern correlation family. 

# Load data and auxiliary functions

The data consists of 

* `pop.dat.km` raster object of the Nigeria population aggregated to 43485 cells, `NA` values were replaced by 0. 

```{r }
load(file.path(lap_data, "Data/pop.RData", fsep = ""))
pop.dat.km = setMinMax(pop.dat.km)
source.all(file.path(lap_data, "Code/auxiliary_functions/", fsep = ""))
```

# Scenarios 

In each scenario for one of the specified models (`iid`, `bym2`, `spde`):

* Number of the replicated data sets is 100.
* The smoothness parameter of the spatial field is fixed, because we vary the standard error $\sigma$ of the field. 
* Using function `scenario_param`, we create 16 scenarios based on the four parameters.
* The possible values of the parameters are in the brackets. The parameters are: spatial fields' range ($7.8\%$ or $38\%$ of the domain extent); spatial fields' variance ($0.2$ or $0.4$); prevalence ($0.20 - 0.30 \%$ or $1.45 - 1.60 \%$ of the 180 mln population) encoded as an intercept $\beta_{0}$ ($-6$ or $-4.3$) and administrative level (level 1 with 37 states or level 2 with 773 Local Government Areas (LGA)).
* The boundaries of the spatial polygon `map.km` are used to determine the values for the range parameters. 

```{r warning=FALSE}
map <- st_geometry(gb_adm1("nigeria"))  
n = length(map)
map <- as_Spatial(map, IDs =as.character(1:n)) 
# Transform the CRS to km 
map.km <- spTransform(
  map, CRS('+proj=utm +zone=23 +south +ellps=GRS80 +units=km'))

scenarios = scenario_param(sp = map.km)

sum.scen = scenarios[[1]]
for(j in 2:16){
 sum.scen = rbind(sum.scen, scenarios[[j]])
}
kable(sum.scen, caption = "Summary of all scenarios")


scenario.idx = 1
model = "bym2" # iid, spde
weights = "pop.int" 
n.rep = 1
range = scenarios[[scenario.idx]]$range
sigma2 = scenarios[[scenario.idx]]$sigma2u

beta0 = scenarios[[scenario.idx]]$beta0
adm.lev = scenarios[[scenario.idx]]$adm.lev

```

## Nigeria administrative boundaries

The Nigeria administrative boundaries were obtained from package \texttt{rgeoboundaries}, which can be found on the \href{https://rdrr.io/github/dickoa/rgeoboundaries/}{website of the GitHub package}.

* `map.km`

```{r warning=FALSE}
if(adm.lev == 1){
 map <- st_geometry(gb_adm1("nigeria"))  
} else {
 map <- st_geometry(gb_adm2("nigeria"))  
}

n = length(map)
map <- as_Spatial(map, IDs =as.character(1:n)) 
# Transform the CRS to km 
map.km <- spTransform(
  map, CRS('+proj=utm +zone=23 +south +ellps=GRS80 +units=km'))
plot(map.km)
```

# Algorythm of the data simulation

Algorithm:

1. The raster `pop.dat.km` contains the population values $n_{k}$ of Nigeria at the fine-scale level, $k = 1, \ldots, m$.

2. Let `coo` denote the spatial coordinates of the population's raster cells. Simulate the underlying Gaussian Field at $m$ locations `coo`. We assume the spatially structured random effect follows a zero-mean Gaussian process with the Matérn correlation function. 
We use the pre-specified parameters for range and spatial field variance. 

3. Simulate the linear predictor $\eta_{k} = \operatorname{logit}(p_{k})$, where $p_{k}$ denote the incidence proportion (or 'reference' specific risk) at the kth raster cell. 
Linear predictor is additive with respect to other effects
\begin{align}
 \eta_{k} &= \beta_{0} +  b_{k},
\end{align} 
where $\beta_{0}$ is an intercept, quantifying the average outcome rate in the study region, $b_{k}$ a spatially structured random effect.
We use the pre-specified parameter for the intercept.  For example $\beta_{0}=-6$ gives 
prevalence $0.20 - 0.30 \%$ of the 180 mln population. 

3. Create the risk surface. 
The relative risk value is computed as 
\begin{align}
r_{k} &= \frac{p_{k}}{\bar{p}} \\
\bar{p} &= \frac{\sum_{k}p_{k}n_{k}}{\sum_{k}n_{k}}.
\end{align}

4. Simulate the disease counts.
Let $Y_{k}$ denote the random variable at the $k$th cell, such that 
\begin{align}
Y_{k}|\eta_{k} &\sim \operatorname{Poisson}(n_{k}p_{k} = n_{k} \frac{e^{\eta_{k}}}{1+e^{\eta_{k}}}). \\
\end{align}

5. Aggregate the population and disease count values to the administrative levels.  
We consider two administrative levels with 37 or 773 areas. 

6. Compute the expected counts for each area $i$, $i = 1, \ldots, n$ 
$$
E_{i} = n_{i} \frac{\sum_{i}y_{i}}{\sum_{i}n_{i}},
$$
which implies 
$$
\sum_{i} E_{i} = \sum_{i} y_{i}. 
$$

# 1. Population of Nigeria

The population of Nigeria was obtained from the
\href{https://apps.worldpop.org/woprVision/}{website of World population}. 

```{r}
# Raster is loaded 
pop.dat.ng <- crop(pop.dat.km, map.km)
plot(pop.dat.ng)
```
## Combining map with data frame 

* The `df` data frame contains `coordinates` of the Nigeria map, that are labelled as `ID`.  

```{r}
# Create a SpatialPolygonsDataFrame 
df = data.frame(lat = coordinates(map.km)[,1], 
                lon = coordinates(map.km)[,2],  
                 ID =  seq(1:length(map.km)),
                 row.names = 1:length(map.km))
# Polygons ids are ID i
mapdf.km = SpatialPolygonsDataFrame(map.km, df) 
```


# 2. The Gaussian random field 

[@abrahamsen1997review] review Gaussian random fields and correlation functions.
A Gaussian random field (GF) is a stochastic process, $\{B(\pmb{s}): \pmb{s} \in D \subset \mathbb{R}^2 \}$. 

We assume the Gaussian field $B(\pmb{s})$ is

* continuous over space $D$. Continiouity implies, that it is possible to collect its realization at any finite set of locations.
* stationary and isotropic, i.e., its covariance function depends only on the distance $h$ and it does not depend on the direction. 

The realization of the GF, at finite set of locations $(b(\pmb{s}_{1}), \ldots, b(\pmb{s}_{m}))$ has a stationary multivariate Gaussian distribution, i.e $$(b(\pmb{s}_{1}), \ldots, b(\pmb{s}_{m})) \sim \operatorname{MVN}(0, \pmb{\Sigma}).$$  


## Matern cavariance function 

[@guttorp2006studies] present a history of the Matern family of spatial correlations. 

Matern covariance function is of the form 
$$
\Sigma_{ij} = C(h) = \sigma^2 2^{1- \nu} \Gamma(\nu)^{-1} (\kappa h)^{\nu} K_{\nu}(\kappa h)
$$

where

* $\kappa$ is a scale parameter,
* $\nu$ is a smoothness parameter,
* $\sigma^2$ is the marginal variance of the process,
* $h = ||\pmb{s}_{i} - \pmb{s}_{j}||$ denote the Euclidian distance betweeen two spatial points,
* $K_{\nu}$ is the modified Bessel function of the second kind.

## Simulation of the GF 

Using the `RandomFields` package, the simulation of the GF is computationally efficient. 
See `RMwhittlematern`, for the parametrization of the Matern correlation function. 
We use the scale parameter equal to half of the range, where range is distance that gives the correlation near 0.14 and has the expression $$ \operatorname{range}= \frac{\sqrt{8 \nu}}{\kappa} $$

```{r warning=F}
coo = xyFromCell(pop.dat.km, 1:ncell(pop.dat.km)) # coordinates from all cells of raster r
colnames(coo) = c("coords.x1", "coords.x2")

m.model <- RMmatern(nu = 1, var = sigma2, scale = range/2)
u <-  RFsimulate(m.model, x= coo[,1], y= coo[,2])


r.GF = pop.dat.km # copy the raster but change the values 
values(r.GF) <- u@data$variable1 #  assign the data values to the raster object
plot(r.GF)
```

# 3 - 7 Simulation of the coherent generative model 

Simulate the risk surface at the fine-scale level, observed and the expected disease counts at the areal level. 

```{r}
  # 3. linear predictor
  eta = beta0 + values(r.GF) 
  
  # 4. relative risk surface
  p.true = exp(eta)/(1+exp(eta)) 
  pop = values(pop.dat.km)
  p.av.true = sum(pop*p.true)/sum(pop) # average probability
  rr.true = p.true/p.av.true # relative risk 
  
  # 5. disease counts based on true p and n
  ysim = rpois(dim(coo)[1], lambda = pop*p.true) 
  r.y <- pop.dat.km
  values(r.y) <- ysim

  # 6. pop and counts aggregation
  mapdf.km@data$pop <- extract(pop.dat.km, map.km, fun = sum, na.rm = TRUE) 
  mapdf.km@data$count <- extract(r.y, map.km, fun = sum, na.rm = TRUE) 
  
  # 7. expected counts
  E = mapdf.km@data$pop*sum(mapdf.km@data$count)/sum(mapdf.km@data$pop) 
  mapdf.km@data$E <- E 

```

## Visualization 

```{r }
r.RRt = pop.dat.km
values(r.RRt) = rr.true
r.RRt <- mask(r.RRt, map.km)
plot(r.RRt, main = c("Relative Risk"))
```

# Inference with INLA - iid, BYM2, SPDE 

For a given scenario, we perform the inference for one of the specified models: `iid`, `bym2`, or `spde`. 


## Precision matrix for BYM2

A spatial adjacency matrix is required for `bym2` model fitting. We create the graph and adjacency matrix based on the neighboring structure encoded in the spatial polygon `map.km`. The graph `ng.graph` is defined once for all simulations.

```{r}
# create rook (right) adjacency matrix, the simplest, condiering neighbours
adj.r <- poly2nb(map.km, queen = FALSE) # not a queen
# convert nb object to graph object compatible with INLA, save it and read it
nb2INLA(file.path(lap_data, "graph", eval(scenario.idx), eval(model), fsep = ""), adj.r)
ng.graph = inla.read.graph(file.path(lap_data, "graph", eval(scenario.idx), eval(model), fsep = ""))

image(inla.graph2matrix(ng.graph), xlab = "Node index", ylab = "Node index")
```

## Mesh, SPDE model and projector matrix 

* If we use the `spde` model, we need to define triangulation of the domain `mesh`, `spde` model with the prior specification, projector matrix for areal data, `Aa` and prediction lattice and matrix.  These are defined once for all simulations. 
* The `mesh` is defined based on the spatial polygon `map.km`
* The `spde` model priors are defined based on true value $\sigma \in \{ 0.2, 0.4\}$ and the true value of range$\in \{121, 600 \}$.
* Projector matrix `Aa` is created based on the spatial polygons `map.km` and the mesh nodes.

```{r}
# mesh
param = areal_mesh_param(spdf = map.km)
mesh = create_mesh(spdf = map.km, max.edge = param$max.edge, cutoff = param$cutoff,
                   offset = param$offset, plot = T)
# spde model
spde <- inla.spde2.pcmatern(
    mesh, alpha=2,
    prior.sigma=c(1, 0.01),  # P(sigma > 1) = 0.01, 
    prior.range=c(range, 0.5)) # P(r < range) = 0.5
  
# projector matrix
  if(weights == "equal" ){
  Aa <- create_matrix(sp = map.km, mesh = mesh, weights = "equal")
  } else {
  Aa <- create_matrix(sp = map.km, mesh = mesh, weights = "pop.int", pop.raster = pop.dat.ng)
  }
  
# prediction lattice and matrix 
# coo are the coordinates of the population raster
Ap = inla.spde.make.A(mesh = mesh, loc = coo)
  
plot(mapdf.km) # region of interest 
points(coo, pch = 3, cex = .01) # prediction points
points(coordinates(map.km), cex = 0.5, pch = 4, col = "blue") # observed points

```


## Fitting - SPDE, BYM2

* Stack and spde model are required for the spde model fitting. 
* Due to the computational efficiency, the Monte Carlo samples are used to predict the continuous risk surface at the population's raster coordinates. 
* Neighbouring structure graph is required for the BYM2 model fitting. The predictions of the risk surface are obtained at the administrative boundaries level.

```{r }
# SPDE
stack.a <- create_stack(tag = "area", matrix = Aa, spdf = mapdf.km, spde = spde)
res <- fit_spde(spde = spde, stack = stack.a)
rrMC = MC_samples(result = res, N = 2e3, Ap = Ap)

# BYM2
formula.model <- count ~ 1 + f(ID, model = "bym2", graph = ng.graph,
                       constr = T, scale.model = T, 
                       hyper=list(prec=list(prior="pc.prec", param=c(1, 0.01)), # P(sd >1) = 0.01
                                  phi =list(prior="pc", param=c(0.5, 0.5)))) # P(phi <0.5) = 0.5
res = inla(formula = formula.model, 
           family = "poisson", data = mapdf.km@data, E =  E, 
           control.compute = list(dic = T, waic = T, cpo = T, config = TRUE),
           control.predictor = list(compute = TRUE),
           num.threads="4:1", quantiles = c(0.1, 0.9))

mapdf.km@data$RR.m <- res$summary.fitted.values[, "mean"]
mapdf.km@data$RR.qu <- res$summary.fitted.values[, "0.9quant"]
mapdf.km@data$RR.ql <- res$summary.fitted.values[, "0.1quant"]

```

## Relative risk mapping - SPDE, BYM2

```{r}
# visualize the spde prediction
r.RRm <- r.RRt
values(r.RRm) <- rrMC$rr$mean
r.RRm = mask(r.RRm, map.km)
plot(r.RRm, main = "Posterior RR mean - SPDE")
lines(map.km)

# visualize the bym2 prediction
r.RRm <- rasterize(mapdf.km, pop.dat.km, field = "RR.m", fun = "sum") 
# fun "sum"" or default "last" gives the same value beacuse we rasterize the polygons
r.RRm <- mask(r.RRm, map.km) # bc of NA
plot(r.RRm, main = "Posterior RR mean - BYM2") 
lines(map.km)

```

# Simulation study 

In the simulation study, we replicate the dataset `n.rep` number of times and compare the models based on the model performance measures. 

# Model performance measures

Let 
$$
\pmb{r} = (r_{1}, \ldots, r_{m})
$$ 

denote the vector of the relative risk at the high resolution grid inside the spatial polygon of interest, where $m$ is the number of cells. 
Let $\hat{r}_{k}^{j}$ denote the posterior mean of the fitted relative risk at $k$th cell and $j$th simulation, where $k = 1, \ldots, m $ and $j =1, \ldots, 100$. For the iid or bym2 model, this estimate has the same value for all cells that fall in the  area. The absolute error, which depends on $j$th simulation is

$$
\operatorname{AE}(r_{k}^{j}) = |\hat{r}_{k}^{j}-r_{k}|
$$

The overall mean absolute error for the relative risk is computed as

$$
\operatorname{MAE}(\pmb{r}) = \frac{1}{100}\sum_{j}^{100}\frac{1}{m}\sum_{k}^{m} \operatorname{AE}(r_{k}^{j})
$$

In a given simulation $j$ 

$$
\operatorname{coverage}(r_{k}) = \begin{cases}
1 & \quad r_{k} \in (q_{l}(\hat{r}^{j}_{k}), q_{u}(\hat{r}^{j}_{k}))\\
0 & \quad \text{ow.}\\ 
\end{cases} 
$$

The overall coverage is computed as

$$
    \operatorname{coverage}(\pmb{r}) = \frac{1}{100} \sum_{j=1}^{100}
\frac{1}{m} \sum_{k=1}^{m} \operatorname{coverage}(r_{k}).
$$

Additionally, we compute the credible interval lengths and their mean across cells and simulations.

```{r}
# globally defined 
# m.model, coo, r.GF, pop, r.RRt, r.y, mapdf.km@data$pop pop.dat.km, map.km, r.RRm

sim.coverage = function(id) {
  
  ## simulate the data 
  u <-  RFsimulate(m.model, x= coo[,1], y= coo[,2]) #  update the GF values
  values(r.GF) <- u@data$variable1
  # 3. linear predictor
  eta = beta0 + values(r.GF) 
  # 4. relative risk surface
  p.true = exp(eta)/(1+exp(eta)) 
  p.av.true = sum(pop*p.true)/sum(pop) # average probability
  rr.true = p.true/p.av.true # relative risk 
  values(r.RRt) = rr.true # update the simulated relative risk values
  r.RRt <- mask(r.RRt, map.km) # contains NA
  # 5. disease counts based on true p and n
  ysim = rpois(dim(coo)[1], lambda = pop*p.true) 
  values(r.y) <- ysim
  # 6. pop and counts aggregation
  mapdf.km@data$count <- extract(r.y, map.km, fun = sum, na.rm = TRUE) 
  # 7. expected counts
  E = mapdf.km@data$pop*sum(mapdf.km@data$count)/sum(mapdf.km@data$pop) 
  mapdf.km@data$E <- E 

  ## INFERENCE and prediction
  if(model == "spde"){
  stack.a <- create_stack(tag = "area", matrix = Aa, spdf = mapdf.km, spde = spde)
  res <- fit_spde(spde = spde, stack = stack.a)
  # prediction and CI 
  rrMC = MC_samples(result = res, N = 2e3, Ap = Ap)
  r.RRm <- r.RRt
  values(r.RRm) <- rrMC$rr$mean
  r.RRm = mask(r.RRm, map.km) # contains NA
  
  r.RRql = r.RRqu = r.RRm
  values(r.RRql) = rrMC$rr$q1
  r.RRql = mask(r.RRql, map.km)
  values(r.RRqu) = rrMC$rr$q9
  r.RRqu = mask(r.RRqu, map.km)
  CI <- cbind(values(r.RRql), values(r.RRqu))
  } else {
  if(model == "bym2"){
  formula.model <- count ~ 1 + f(ID, model = "bym2", graph = ng.graph,
                       constr = T, scale.model = T,
                       hyper=list(prec=list(prior="pc.prec", param=c(1, 0.01)), # P(sd >1) = 0.01
                                  phi =list(prior="pc", param=c(0.5, 0.5)))) # P(phi <0.5) = 0.5) 
 } else {
  formula.model <- count ~ 1 + f(ID, model = "iid")
 }
  res = inla(formula = formula.model, 
           family = "poisson",
           data = mapdf.km@data, 
           E =  E, 
           control.compute = list(dic = T, waic = T, cpo = T, config = TRUE),
           control.predictor = list(compute = TRUE),
           num.threads="4:1",
           quantiles = c(0.1, 0.9))
  # prediction and CI 
  mapdf.km@data$RR.m <- res$summary.fitted.values[, "mean"]
  mapdf.km@data$RR.qu <- res$summary.fitted.values[, "0.9quant"]
  mapdf.km@data$RR.ql <- res$summary.fitted.values[, "0.1quant"]
  r.RRm <- rasterize(mapdf.km, pop.dat.km, field = "RR.m", fun = "sum") # has NA
  r.RRql <- rasterize(mapdf.km, pop.dat.km, field = "RR.ql", fun = "sum") # CI
  r.RRqu <- rasterize(mapdf.km, pop.dat.km, field = "RR.qu", fun = "sum") # CI 
  CI <- cbind(values(r.RRql), values(r.RRqu))
  }
  
  ## model performance measures
  # absolute error
  ae = abs(values(r.RRt) - values(r.RRm)) # true - posterior mean 
  # credible interval lenght
  cov.len = CI[,2] - CI[,1]
  # binary coverage vector 
  cov.bin <- c()
  for(idx in 1:length(r.RRt)){
   cov.bin[idx] = between(values(r.RRt)[idx], CI[idx, 1], CI[idx, 2])
  }

  # save the perf measures for cells inside spatial polygons
  return(list(cov.bin = cov.bin[!is.na(cov.bin)], 
             cov.len = cov.len[!is.na(cov.len)], 
             ae = ae[!is.na(ae)]))

}

```

Results for one simulation in a given scenario.

```{r}
perf = sim.coverage(id = 1)
mean(perf$cov.bin)
mean(perf$ae)
mean(perf$cov.len)
```

# Computations on the server 

The simulations were done on the server, running 64 tasks in parallel and using 125 GB memory.
Fitting 46 scenarios took approximately X hours.
The code below is presented, but not executed. 
The obtained list of lists `res.sim.raw` was processed. 
Each list (for a given simulation) consited of `cov.bin`, `cov.len` and `ae` vectors of size $m$.

For each scenario, we saved the matrices `cov.bin`, `cov.len` and `ae` in `mp.Rdata` files, which are processed in the next subsection.

```{r eval = F}
## Run the model for n.rep number of datasets
res.sim.raw = mclapply(as.list(1:n.rep), mc.cores=3, FUN = sim.coverage)
m = length(res.sim.raw[[1]]$cov.bin)

## Extract model performance matrices of dimension n.rep x m
cov.bin <- matrix(NA, nrow = n.rep, ncol = m) 
cov.len <- matrix(NA, nrow = n.rep, ncol = m) 
ae <- matrix(NA, nrow = n.rep, ncol = m) 

for(id in 1:n.rep){
  # index of the list corresponds to the sim index
  cov.bin[id, ] = res.sim.raw[[id]]$cov.bin
  cov.len[id, ] = res.sim.raw[[id]]$cov.len
  ae[id, ] = res.sim.raw[[id]]$ae
}

## Save the results 
save(cov.bin, cov.len, ae, file = file.path( "Results/mp", eval(scenario.idx), 
                                             eval(model), ".RData", fsep = "")) 

```

## Server computaion details 

The script without plots and text was run using the terminal:

* Log in to the server using the `ssh server` command.
* Execute the *shell* scripts using `./run.jobs` command 
where the scripts `run.jobs` is as follows

```{r eval =F}
#!/bin/bash -x
for sc in $(seq 1 16); do
    for lik in "spde" "bym2" "iid"; do
        nice Rscript simraw.R $sc $lik  > LOG-$sc-$lik 2>&1 &
    done
done
```

* To be able to execute the `simraw.R` script over the scenario and model arguments, we need to define the `input` at the beginning of the `simraw.R` script:

```{r, eval=F}
input = commandArgs(trailingOnly=TRUE)
scenario.idx = as.integer(input[1])
model = input[2]
```

* The `input[1]` will take the value of one of the 16 scenarios, the `input[2]` will take the value of one of the three models.
* Both `inputs` are passed to the script.  Hence, we can run the scripts in parallel for different input combinations. 
* The results are saved depending on the scenario and the model. 

# Processing the results - read the results 

* We read the results, that were saved from the runs on the server. 
* For each of the 48 scenarios, we obtained the model performance measures: credible interval length, binary vector if the true value falls in CI, and absolute error. 
* For a given scenario and for all 100 replicate datasets, we have three matrices of size $100 \times m,$ where $m$ is a number of raster cells inside the spatial polygon of interests. 
* We create `cov.bin.s`, `cov.len.s` and `ae.s` lists to store the matrices from all scenarios. 
 

```{r}
cov.bin.s = list()
cov.len.s = list()
ae.s = list()
combination = list()
idx = 1
set = c(seq(1:10), 13,14,15) # 16 

for(scenario.idx in set){
  for( model in c("iid", "bym2", "spde")){ 
   ## Save the results 
   load(file.path(lap_data, "Results/mp", eval(scenario.idx),eval(model), ".RData", fsep = "")) 
   cov.bin.s[[idx]]  = cov.bin
   cov.len.s[[idx]]  = cov.len
   ae.s[[idx]]  = ae
   combination[[idx]] = paste(scenario.idx, model)
   idx = idx +1
  }
}

```

# Boxplot of the absolute error 

Let `n.mod` denotes the number of the models considered, `m` number of raster cells inside the spatial polygon of interests, and `n.rep` number of replicated data sets. 
We create a boxplot of the absolute error obtained for $m$ cells across $n.rep$ simulations. 
We present boxplots for three analyzed models together. We create two different plots for all scenarios. 

To do so, we create a data frame consisting of 3 columns:

* The first column indicate `scenario` (from 1 to `n.sc`). Each `scenario` index is
repeated `n.mod` $\times$ `m` $\times$  `n.rep` number of times.
* The second column is a vector denoting the model name. Each model name is repeated `m` $\times$  `n.rep` number of times per each scenario. 
* For a given scenario and a given model, we obtain a vector (vectorized matrix) of the absolute errors of size `m` $\times$  `n.rep`.

```{r}
n.sc = 13 # 16
n.mod = 3
n.rep = dim(ae.s[[1]])[1]
m = dim(ae.s[[1]])[2]

## create a data frame
large.ae = c() 
# define the scenario names
sc.names = c("LRLVLP1", "LRLVHP1","LRHVLP1", "LRHVHP1", 
             "HRLVLP1", "HRLVHP1","HRHVLP1", "HRHVHP1",
             "LRLVLP2", "LRLVHP2","LRHVLP2", "LRHVHP2", 
             "HRLVLP2", "HRLVHP2","HRHVLP2", "HRHVHP2")

dt = data.frame(scenario = as.character(rep(sc.names[c(seq(1:10), 13,14,15)], 
                                            each = n.mod*m*n.rep)),
                  model = rep(c("iid", "bym2", "spde"), each = m*n.rep))
for(i in 1:length(combination)){
  large.ae = c(large.ae,as.vector(as.matrix(ae.s[[i]])))
}
dt$ae = large.ae


# split scenarios in two groups by name
# set = list(g1 = sc.names[1:8], g2 = sc.names[9:16]) 
set = list(g1 = sc.names[1:8], g2 = sc.names[c(9,10, 13,14,15)]) 
s = 1
p <- list()

# to save the plots at the selected path do not save them under list p
for(v in set){
  p[[s]] <- ggplot(dt[dt[,"scenario"] %in% v,], 
          aes(x = scenario, y = ae, fill = model)) + geom_boxplot() +
          facet_wrap(~scenario, scale = "free", ncol = 4) + theme_bw() + ylab("Absolute error") +
          xlab("Scenario name") + ylim(c(0, 3))
  # ggsave(file.path(lap_data, "ae", names(set)[s], ".jpeg", fsep = ""))
  s = s+1
}
p[[1]]
p[[2]]
```

# Model performace summary for all scenarios

The comparison is done overall row and column - wise average of `n.rep` $\times$  `m`.

```{r }
ss = matrix(NA, nrow = length(combination), ncol = 10)

ss[,1]<- unlist(combination)
ss[,2] <- rep(unlist(sum.scen[c(seq(1:10), 13,14,15), "range"]), each = 3) # we have three models 
ss[,3] <- rep(unlist(sum.scen[c(seq(1:10), 13,14,15), "sigma2u"]), each = 3) 
ss[,4] <- rep(unlist(sum.scen[c(seq(1:10), 13,14,15), "beta0"]), each = 3) 
ss[,5] <- rep(unlist(sum.scen[c(seq(1:10), 13,14,15), "adm.lev"]), each = 3) 

for(i in 1:length(combination)){
  ss[i,6] = round(mean(as.matrix(cov.bin.s[[i]])),3) # overal
  ss[i,7] = round(mean(as.matrix(cov.len.s[[i]])),3) # overal
  ss[i,8] = round(mean(as.matrix(ae.s[[i]]), na.rm = TRUE),3)
  ss[i,9] = round(quantile(as.matrix(ae.s[[i]]),0.025, na.rm = TRUE),3)
  ss[i,10] = round(quantile(as.matrix(ae.s[[i]]),0.975, na.rm = TRUE),3)

}

colnames(ss) <- c("Combination","range","sigma2u", "beta0",
                                "adm.lvl", "Coverage", "CIs lenght", "MAE", "AEq0.05", "AEq0.95")
```

Averages of the $100 \times m$ performance measures across all scenarios are:

```{r }
kable(ss, caption = "Summary results across all models and scenarios")
write.csv(ss,file = file.path(lap_data, "Results/summaryscenario.csv", fsep = ""))
```

# References 

